\vspace{-3em} \textbf{ Mapping stellar surfaces\\ I: Degeneracies in the rotational light curve problem }

### 2021-03-30 paper1/ms.tex
add line numbers
reviewer comments; push dataset to zenodo

For rotational light curves,
the vast majority of the surface modes lie in the null
space. To show this, we will make use of the fact that we can express
the vector of $K$ observed fluxes $\mathbf{f}$ (i.e., the light curve)
as a linear operation on the vector of $N$
spherical harmonic coefficients $\mathbf{y}$ \citep{Luger2019}:
%
\begin{align}
    \label{eq:fAy}
    \mathbf{f} = \mathbf{1} + \pmb{\mathcal{A}} \, \mathbf{y}
    \quad,
\end{align}
%
where $\pmb{\mathcal{A}}$ is the $(K \times N)$ \emph{design matrix} of the transformation, whose columns
describe how each of the $N$ components in the spherical harmonic basis contribute
to each of the $K$ points in the light curve.%
\footnote{%
    For rotational light curves, the rows of $\pmb{\mathcal{A}}$ are given by the
    quantity $\mathbf{r}^\top \pmb{\mathcal{A}}_1 \mathbf{R}$ in Equation~(18) of
    \citet{Luger2019}, where $\mathbf{r}^\top$ is a vector of disk-integrated
    intensities, $\pmb{\mathcal{A}}_1$ is a change of basis matrix, and $\mathbf{R}$
    is a spherical harmonic rotation matrix that depends on the stellar inclination
    and the rotational phase of the star. In Equation~(\ref{eq:fAy}) we explicitly
    add a vector of ones to enforce a unit baseline for spotless stars.
    Refer to \citet{Luger2019} for more details.
}
%
Even though we are explicitly choosing the spherical harmonics as the basis in which
we describe the stellar surface, Equation~(\ref{eq:fAy}) is quite general and applies to
\emph{any} basis that is linearly related to the flux. For instance, $\mathbf{y}$
could instead describe the intensity deficits in each of the $N$ pixels of a gridded stellar
surface, in which case $\pmb{\mathcal{A}}$ would be the matrix of pixel visibilities that describe
how to sum each of the pixels to obtain the observed light curve.

@@

\subsection{Decomposition of the flux operator}
%
\label{sec:svd}
%
In Appendix~\ref{sec:app-svd} we show that it is straightforward to
construct linear operators $\mathbf{P}$ and $\mathbf{N}$ from the
design matrix $\pmb{\mathcal{A}}$ such that
%
\begin{align}
    \mathbf{y}_\bullet & = \mathbf{P} \, \mathbf{y}
\end{align}
%
is the component of the surface map in the \emph{preimage} and
%
\begin{align}
    \mathbf{y}_\circ & = \mathbf{N} \, \mathbf{y}
\end{align}
%
is the component of the map in the \emph{null space}.
%
The \emph{preimage operator} $\mathbf{P}$
transforms a vector $\mathbf{y}$ in the surface map basis in such a way
that it preserves the information in $\mathbf{y}$ that gets mapped
onto the light curve $\mathbf{f}$ via $\pmb{\mathcal{A}}$ (the \emph{preimage}) and discards the rest. The
\emph{null space operator} $\mathbf{N}$ does the opposite: it preserves only
the information in $\mathbf{y}$ that gets mapped onto the zero
vector via $\pmb{\mathcal{A}}$ (the \emph{null space}).
%
In other words,
the $\mathbf{P}$ and $\mathbf{N}$ operators reveal the
components of the surface map that contribute to the
light curve ($\mathbf{y}_\bullet$) and the components that don't ($\mathbf{y}_\circ$).
The vector $\mathbf{y}_\bullet$ represents all the information that can be
learned from a stellar light curve, while $\mathbf{y}_\circ$ represents all the
information that cannot.

@@

\subsection{Implications for inference}
%
The model mapping spherical harmonic coefficients to observed fluxes described
in the preceding sections is linear, so we might expect that we could fit
for the spherical harmonics using linear least-squares
%
\begin{align}
    \label{eq:linlstsq}
    \hat{\mathbf{y}} & \stackrel{?}{=}
    \left(\pmb{\mathcal{A}}^\top\,\pmb{\mathcal{A}}\right)^{-1}\,\pmb{\mathcal{A}}^\top\,\left(\mathbf{f} - \mathbf{1}\right)
    \quad.
\end{align}
%
But, since the flux operator $\pmb{\mathcal{A}}$ is low-rank, the model is
underdetermined, so the above operation is not defined and, equivalently,
the Fisher information matrix
%
\begin{align}
    \label{eq:fisher-info}
    \pmb{\mathcal{I}}(\mathbf{y}) & = \frac{1}{\sigma_f^2} \, \pmb{\mathcal{A}}^\top\,\pmb{\mathcal{A}}
\end{align}
%
is singular.
In Equation~(\ref{eq:fisher-info}), $\sigma_f$ is the flux measurement uncertainty.
There exist many methods and procedures for linear regression for underdetermined
models, but all solutions amount to imposing stronger assumptions about the
model in order to break the degeneracies.

One standard method is regularized least squares, in which
Equation~(\ref{eq:linlstsq}) becomes
%
\begin{align}
    \hat{\mathbf{y}}_\lambda & \equiv
    \left(\pmb{\mathcal{A}}^\top\,\pmb{\mathcal{A}} + \lambda\,\mathbf{I}\right)^{-1}\,\pmb{\mathcal{A}}^\top\,\left(\mathbf{f} - \mathbf{1}\right)
\end{align}
%
where $\mathbf{I}$ is the identity matrix and $\lambda$ is a parameter
controlling the strength of the regularization.
In a Bayesian context, this can be interpreted as placing a Gaussian prior
with variance $\sigma_0^{2} = \nicefrac{\sigma_f^2}{\lambda}$ on the spherical harmonic coefficients.

In the limit of infinitesimal regularization, it can be demonstrated
\citep[see][for example]{Hogg2021} that
%
\begin{proof}{test_svd_lstsq}
    \label{eq:inference}
    \lim_{\lambda \rightarrow 0_+} \hat{\mathbf{y}}_\lambda
    % &=
    % \lim_{\lambda \rightarrow 0} \left( \pmb{\mathcal{A}}^\top \pmb{\mathcal{A}} + \lambda \mathbf{I}\right)^{-1} \pmb{\mathcal{A}}^\top (\mathbf{f} - \mathbf{1})
    % \nonumber\\
    &= \mathbf{y}_\bullet
    \quad.
\end{proof}
%
In this sense, $\mathbf{y}_\bullet$ represents our knowledge about the
surface of a star after an observation if we have no prior information
whatsoever on $\mathbf{y}$.
\emph{Any other information about the surface is entirely driven by our assumptions.}%

% It is interesting to quantify the information content of a light curve as a
% function of spherical harmonic coefficients.
% While Equation~(\ref{eq:fisher-info}) provides a compact expression for the
% Fisher information matrix, the most important feature that we want to quantify
% is the covariance between coefficients.
% This is most naturally captured by the covariance matrix of the likelihood
% function for the spherical harmonic coefficients:
% %
% \begin{align}
%     \pmb{\Sigma}_\lambda & =
%     \sigma^2\,\left(\pmb{\mathcal{A}}^\top\,\pmb{\mathcal{A}} + \lambda\,\mathbf{I}\right)^{-1} \quad.
% \end{align}
% %
% Below, we use this to quantify the width of the likelihood for each spherical
% harmonic coefficient by plotting the ``relative variance'' $S$, which we define
% as the diagonal of the matrix $\pmb{\Sigma}_\lambda$ in units of
% $\lambda / \sigma^2$ in the limit where $\lambda \to 0_+$.
% This choice of units constrains $0 \le S \le 1$ and allows easy comparisons
% between the information content of different datasets without being sensitive
% to the specific modeling assumptions.
% Under this definition, when $S$ approaches one for a particular coefficient,
% this means that the likelihood function provides no information about that
% coefficient, while $S = 0$ indicates that that coefficient can be perfectly measured.

@@

To better understand this effect, let us define the quantity
%
\begin{align}
    S \equiv
    1 -
    \lim_{\lambda \rightarrow 0_+}
    \mathrm{diag}
    \left(
    \lambda
    \left(\pmb{\mathcal{A}}^\top\,\pmb{\mathcal{A}} + \lambda\,\mathbf{I}\right)^{-1}
    \right)
    \quad,
\end{align}
%
which we will refer to as the \emph{\shrinkage}
of the coefficients $\mathbf{y}$ characterizing the stellar surface.
Note that in a Bayesian context, this is
equivalent to the \emph{posterior shrinkage}
%
\begin{align}
    \label{eq:shrinkage}
    S = 1 - \lim\limits_{\sigma_0^2 \rightarrow \infty}
    \frac{\sigma^2}{\sigma_0^2}
    \quad,
\end{align}
%
where $\sigma_0^2$ is the prior variance
and $\sigma^2$ is the posterior variance
on a particular surface mode we are trying to constrain
\citep[see, e.g.,][]{Betancourt2018}.

@@

\begin{figure}[t!]
    \begin{centering}
        \includegraphics[width=\linewidth]{figures/nullspace_ensemble_single.pdf}
        \oscaption{nullspace_ensemble_single}{%
            The \shrinkage as a function of spherical harmonic
            degree $l$ given a single observation of a star at a random
            inclination (thin blue curves). The mean \shrinkage is shown
            as the thicker curve. The information content of light curves
            tends to zero as $l$ increases, and odd $l > 1$ modes are in the
            null space at all inclinations.
            \label{fig:nullspace_ensemble_single}
        }
    \end{centering}
\end{figure}

@@

\begin{figure}[p!]
    \begin{centering}
        %
        %
        \vspace{-4em}
        \includegraphics[width=1in]{figures/wireframe_30.pdf}
        \includegraphics[width=1in]{figures/wireframe_45.pdf}
        \includegraphics[width=1in]{figures/wireframe_60.pdf}
        \includegraphics[width=1in]{figures/wireframe_75.pdf}
        \\[0.5em]
        \includegraphics[width=\linewidth]{figures/nullspace_preimage_a.pdf}
        \\[1em]
        \includegraphics[width=\linewidth]{figures/nullspace_preimage_b.pdf}
        \\[1em]
        \includegraphics[width=\linewidth]{figures/nullspace_preimage_c.pdf}
        \\[1em]
        \includegraphics[width=\linewidth]{figures/nullspace_preimage_d.pdf}
        \oscaption{nullspace_preimage_ensemble}{%
            Same as Figures~\ref{fig:nullspace_preimage} and \ref{fig:nullspace_preimage_85}, \
            but assuming we can measure the light curves of these stars from many
            different inclinations. In this limit, the information content of our
            data approaches $50\%$ of the spatial information about the surface.
            \label{fig:nullspace_preimage_ensemble}
        }
    \end{centering}
\end{figure}

@@

\subsection{Ensemble analyses don't necessarily help}
\label{sec:basic-gp}
%
To explore this point in a bit more detail, let us consider the normalization
degeneracy in the context of ensemble analyses. Even though we can't
uniquely infer contrasts or numbers of spots from individual light curves,
perhaps we could harness the power of the ensemble.
%
Let us therefore go
back to our thought experiment in which we added spots to a stellar
surface. Assuming for simplicity that all spots have the same contrast $c$,
every time we add a spot the flux (in fractional units)
decreases by an amount proportional to $c$, so to first order we can approximate an
individual light curve as
%
\begin{align}
    \label{eq:fapprox}
    \mathbf{f}(c, n) = \mathbf{1} - c \sum_{i=0}^{n-1} \mathbf{g}(\pmb{\theta}_i)
\end{align}
%
where $\mathbf{g}(\pmb{\theta}_i)$ is some (complicated) function of the properties
of the $i^\mathrm{th}$ spot
as well as the stellar inclination, rotation period, etc., which we
denote by $\pmb{\theta}_i$.
Now, consider many stellar light curves drawn from some distribution
controlling the stellar and starspot properties with probability density
$p(\pmb{\theta})$.
%
The mean of the distribution of
light curves (still in fractional units) is then given by
%
\begin{align}
    \mu(c, n) & = \mathrm{E} \Big[ \mathbf{f} (c, n) \Big] \nonumber                                      \\
              & = 1 - c \, \mathrm{E} \bigg[ \sum_{i=0}^{n-1} \mathbf{g}(\pmb{\theta}_i) \bigg] \nonumber \\
              & = 1 - c \, n \, \alpha
    \quad,
\end{align}
%
where $\mathrm{E}\big[\cdots\big]$ denotes the expected value and
%
\begin{align}
    \alpha \equiv \int \mathbf{g} (\pmb{\theta}) p(\pmb{\theta}) \mathrm{d} \pmb{\theta}
\end{align}
%
is the expected value of $\mathbf{g}$.
%
Similarly, the variance of the distribution may be computed as
%
\begin{align}
    \sigma^2(c, n) & = \mathrm{Var} \Big[ \mathbf{f}(c, n) \Big] \nonumber                         \nonumber \\
                   & = \mathrm{Var} \Big[ c \, \sum_{i=0}^{n-1} \mathbf{g}(\pmb{\theta}_i) \Big]   \nonumber \\
                   & = c^2 \, \mathrm{Var} \Big[ \sum_{i=0}^{n-1} \mathbf{g}(\pmb{\theta}_i) \Big] \nonumber \\
                   & = c^2 \sum_{i=0}^{n-1} \mathrm{Var} \Big[  \mathbf{g}(\pmb{\theta}_i) \Big]   \nonumber \\
                   & = c^2 \, n \, (\beta^2 - \alpha^2)
    \quad,
\end{align}
%
where $\mathrm{Var}\big[\cdots\big]$ denotes the variance,
%
\begin{align}
    \beta^2 \equiv \int \mathbf{g}^2 (\pmb{\theta}) p(\pmb{\theta}) \mathrm{d} \pmb{\theta}
\end{align}
%
is the expected value of $\mathbf{g}^2$, and we used the fact that the variance
of the sum of independent random variables is equal to the sum of their variances.

@@

To summarize, the mean and variance of the ensemble of stellar light curves
in fractional units is
%
\begin{align}
    \mu      & = 1 - c \, n \, \alpha
    \nonumber                                     \\
    \sigma^2 & = c^2 \, n \, (\beta^2 - \alpha^2)
    \quad,
\end{align}
%
for some complicated functions $\alpha$ and $\beta$ of the distribution of
stellar inclinations,
rotation periods, and starspot properties.
%
If our observations were collected in these fractional units,
we could uniquely infer the spot contrast $c$ and the number of spots $n$,
since these enter as $c \, n$ and $c^2 n$ in the expressions for the mean
and variance, respectively, and these are straightforward statistics to
compute from the ensemble.
%
However, because of observations are made in \emph{relative} units, in which
we typically normalize to the mean, the amplitudes of features in light curves
can only tell us about the \emph{ratio}
%
\begin{align}
    \label{eq:ratio}
    \frac{\sigma}{\mu}
     & \propto \frac{c \sqrt{n}}{1 - c n \alpha}
\end{align}
%
for some value of $\alpha$.
In other words, even photometric ensemble analyses may not be able to tell
us about the values of the contrast and the number of spots independently.

A direct consequence of this normalization degeneracy is that it may not be
possible to uniquely constrain the total spot coverage of a star from
single-band photometry without strong prior assumptions.
The total spot coverage $f_S$
is simply the (average) area of a spot
times the total number of spots divided by the total area of the sphere, which
may be expressed as
%
\begin{proof}{test_fS}
    \label{eq:fS}
    f_S = \frac{1}{2}\left(1 - \left<\cos r\right>\right)n
    \quad,
\end{proof}
%
given an angular spot radius $r$.
While $r$ may be uniquely constrained from the covariance structure of the data
\citepalias{PaperII}, $n$ cannot.

@@

\begin{proof}{test_decomposition}

@@

\begin{proof}{test_PN}

@@

\begin{proof}{test_orthogonality}

@@

Two things should be noted concerning this derivation. First,
because of the orthogonality of $\mathbf{V}$,
%
\begin{proof}{test_PN}
    \mathbf{V} \mathbf{V}^\top &= \mathbf{I}
    \nonumber                                                 \\
    \mathbf{V}_\bullet \mathbf{V}_\bullet^\top  + \mathbf{V}_\circ \mathbf{V}_\circ^\top &= \mathbf{I}
    \nonumber \\
    \mathbf{P} + \mathbf{N} &= \mathbf{I}
    \quad,
\end{proof}
%
so it follows from Equations~(\ref{eq:yrow}) and (\ref{eq:ynull})
that $\mathbf{y} = \mathbf{y}_\bullet + \mathbf{y}_\circ$.

>>>>

For rotational light curves,
the vast majority of the surface modes lie in the null
space. To show this, we will make use of the fact that we can express
the vector of $K$ observed fluxes $\mathbf{f}$ (i.e., the light curve)
as a linear operation on the vector of $N$
spherical harmonic coefficients $\mathbf{y}$ \citep{Luger2019}:
%
\begin{linenomath}\begin{align}
        \label{eq:fAy}
        \mathbf{f} = \mathbf{1} + \pmb{\mathcal{A}} \, \mathbf{y}
        \quad,
    \end{align}\end{linenomath}
%
where $\pmb{\mathcal{A}}$ is the $(K \times N)$ \emph{design matrix} of the transformation, whose columns
describe how each of the $N$ components in the spherical harmonic basis contribute
to each of the $K$ points in the light curve.%
\footnote{%
    For rotational light curves, the rows of $\pmb{\mathcal{A}}$ are given by the
    quantity $\mathbf{r}^\top \pmb{\mathcal{A}}_1 \mathbf{R}$ in Equation~(18) of
    \citet{Luger2019}, where $\mathbf{r}^\top$ is a vector of disk-integrated
    intensities, $\pmb{\mathcal{A}}_1$ is a change of basis matrix, and $\mathbf{R}$
    is a spherical harmonic rotation matrix that depends on the stellar inclination
    and the rotational phase of the star. In Equation~(\ref{eq:fAy}) we explicitly
    add a vector of ones to enforce a unit baseline for spotless stars.
    Refer to \citet{Luger2019} for more details.
}
%
Even though we are explicitly choosing the spherical harmonics as the basis in which
we describe the stellar surface, Equation~(\ref{eq:fAy}) is quite general and applies to
\emph{any} basis that is linearly related to the flux. For instance, $\mathbf{y}$
could instead describe the intensity deficits in each of the $N$ pixels of a gridded stellar
surface, in which case $\pmb{\mathcal{A}}$ would be the matrix of pixel visibilities that describe
how to sum each of the pixels to obtain the observed light curve.

@@

\subsection{Decomposition of the flux operator}
%
\label{sec:svd}
%
In Appendix~\ref{sec:app-svd} we show that it is straightforward to
construct linear operators $\mathbf{P}$ and $\mathbf{N}$ from the
design matrix $\pmb{\mathcal{A}}$ such that
%
\begin{linenomath}\begin{align}
        \mathbf{y}_\bullet & = \mathbf{P} \, \mathbf{y}
    \end{align}\end{linenomath}
%
is the component of the surface map in the \emph{preimage} and
%
\begin{linenomath}\begin{align}
        \mathbf{y}_\circ & = \mathbf{N} \, \mathbf{y}
    \end{align}\end{linenomath}
%
is the component of the map in the \emph{null space}.
%
The \emph{preimage operator} $\mathbf{P}$
transforms a vector $\mathbf{y}$ in the surface map basis in such a way
that it preserves the information in $\mathbf{y}$ that gets mapped
onto the light curve $\mathbf{f}$ via $\pmb{\mathcal{A}}$ (the \emph{preimage}) and discards the rest. The
\emph{null space operator} $\mathbf{N}$ does the opposite: it preserves only
the information in $\mathbf{y}$ that gets mapped onto the zero
vector via $\pmb{\mathcal{A}}$ (the \emph{null space}).
%
In other words,
the $\mathbf{P}$ and $\mathbf{N}$ operators reveal the
components of the surface map that contribute to the
light curve ($\mathbf{y}_\bullet$) and the components that don't ($\mathbf{y}_\circ$).
The vector $\mathbf{y}_\bullet$ represents all the information that can be
learned from a stellar light curve, while $\mathbf{y}_\circ$ represents all the
information that cannot.

@@

\subsection{Implications for inference}
%
The model mapping spherical harmonic coefficients to observed fluxes described
in the preceding sections is linear, so we might expect that we could fit
for the spherical harmonics using linear least-squares
%
\begin{linenomath}\begin{align}
        \label{eq:linlstsq}
        \hat{\mathbf{y}} & \stackrel{?}{=}
        \left(\pmb{\mathcal{A}}^\top\,\pmb{\mathcal{A}}\right)^{-1}\,\pmb{\mathcal{A}}^\top\,\left(\mathbf{f} - \mathbf{1}\right)
        \quad.
    \end{align}\end{linenomath}
%
But, since the flux operator $\pmb{\mathcal{A}}$ is low-rank, the model is
underdetermined, so the above operation is not defined and, equivalently,
the Fisher information matrix
%
\begin{linenomath}\begin{align}
        \label{eq:fisher-info}
        \pmb{\mathcal{I}}(\mathbf{y}) & = \frac{1}{\sigma_f^2} \, \pmb{\mathcal{A}}^\top\,\pmb{\mathcal{A}}
    \end{align}\end{linenomath}
%
is singular.
In Equation~(\ref{eq:fisher-info}), $\sigma_f$ is the flux measurement uncertainty.
There exist many methods and procedures for linear regression for underdetermined
models, but all solutions amount to imposing stronger assumptions about the
model in order to break the degeneracies.

One standard method is regularized least squares, in which
Equation~(\ref{eq:linlstsq}) becomes
%
\begin{linenomath}\begin{align}
        \hat{\mathbf{y}}_\lambda & \equiv
        \left(\pmb{\mathcal{A}}^\top\,\pmb{\mathcal{A}} + \lambda\,\mathbf{I}\right)^{-1}\,\pmb{\mathcal{A}}^\top\,\left(\mathbf{f} - \mathbf{1}\right)
    \end{align}\end{linenomath}
%
where $\mathbf{I}$ is the identity matrix and $\lambda$ is a parameter
controlling the strength of the regularization.
In a Bayesian context, this can be interpreted as placing a Gaussian prior
with variance $\sigma_0^{2} = \nicefrac{\sigma_f^2}{\lambda}$ on the spherical harmonic coefficients.

In the limit of infinitesimal regularization, it can be demonstrated
\citep[see][for example]{Hogg2021} that
%
\begin{linenomath}\begin{proof}{test_svd_lstsq}
        \label{eq:inference}
        \lim_{\lambda \rightarrow 0_+} \hat{\mathbf{y}}_\lambda
        % &=
        % \lim_{\lambda \rightarrow 0} \left( \pmb{\mathcal{A}}^\top \pmb{\mathcal{A}} + \lambda \mathbf{I}\right)^{-1} \pmb{\mathcal{A}}^\top (\mathbf{f} - \mathbf{1})
        % \nonumber\\
        &= \mathbf{y}_\bullet
        \quad.
    \end{proof}\end{linenomath}
%
In this sense, $\mathbf{y}_\bullet$ represents our knowledge about the
surface of a star after an observation if we have no prior information
whatsoever on $\mathbf{y}$.
\emph{Any other information about the surface is entirely driven by our assumptions.}%

% It is interesting to quantify the information content of a light curve as a
% function of spherical harmonic coefficients.
% While Equation~(\ref{eq:fisher-info}) provides a compact expression for the
% Fisher information matrix, the most important feature that we want to quantify
% is the covariance between coefficients.
% This is most naturally captured by the covariance matrix of the likelihood
% function for the spherical harmonic coefficients:
% %
% \begin{linenomath}\begin{align}
%     \pmb{\Sigma}_\lambda & =
%     \sigma^2\,\left(\pmb{\mathcal{A}}^\top\,\pmb{\mathcal{A}} + \lambda\,\mathbf{I}\right)^{-1} \quad.
% \end{align}\end{linenomath}
% %
% Below, we use this to quantify the width of the likelihood for each spherical
% harmonic coefficient by plotting the ``relative variance'' $S$, which we define
% as the diagonal of the matrix $\pmb{\Sigma}_\lambda$ in units of
% $\lambda / \sigma^2$ in the limit where $\lambda \to 0_+$.
% This choice of units constrains $0 \le S \le 1$ and allows easy comparisons
% between the information content of different datasets without being sensitive
% to the specific modeling assumptions.
% Under this definition, when $S$ approaches one for a particular coefficient,
% this means that the likelihood function provides no information about that
% coefficient, while $S = 0$ indicates that that coefficient can be perfectly measured.

@@

To better understand this effect, let us define the quantity
%
\begin{linenomath}\begin{align}
        S \equiv
        1 -
        \lim_{\lambda \rightarrow 0_+}
        \mathrm{diag}
        \left(
        \lambda
        \left(\pmb{\mathcal{A}}^\top\,\pmb{\mathcal{A}} + \lambda\,\mathbf{I}\right)^{-1}
        \right)
        \quad,
    \end{align}\end{linenomath}
%
which we will refer to as the \emph{\shrinkage}
of the coefficients $\mathbf{y}$ characterizing the stellar surface.
Note that in a Bayesian context, this is
equivalent to the \emph{posterior shrinkage}
%
\begin{linenomath}\begin{align}
        \label{eq:shrinkage}
        S = 1 - \lim\limits_{\sigma_0^2 \rightarrow \infty}
        \frac{\sigma^2}{\sigma_0^2}
        \quad,
    \end{align}\end{linenomath}
%
where $\sigma_0^2$ is the prior variance
and $\sigma^2$ is the posterior variance
on a particular surface mode we are trying to constrain
\citep[see, e.g.,][]{Betancourt2018}.

@@

\begin{figure}[t!]
    \begin{centering}
        \includegraphics[width=\linewidth]{figures/nullspace_ensemble_single.pdf}
        \oscaption{nullspace_ensemble}{%
            The \shrinkage as a function of spherical harmonic
            degree $l$ given a single observation of a star at a random
            inclination (thin blue curves). The mean \shrinkage is shown
            as the thicker curve. The information content of light curves
            tends to zero as $l$ increases, and odd $l > 1$ modes are in the
            null space at all inclinations.
            \label{fig:nullspace_ensemble_single}
        }
    \end{centering}
\end{figure}

@@

\begin{figure}[p!]
    \begin{centering}
        %
        %
        \vspace{-4em}
        \includegraphics[width=1in]{figures/wireframe_30.pdf}
        \includegraphics[width=1in]{figures/wireframe_45.pdf}
        \includegraphics[width=1in]{figures/wireframe_60.pdf}
        \includegraphics[width=1in]{figures/wireframe_75.pdf}
        \\[0.5em]
        \includegraphics[width=\linewidth]{figures/nullspace_preimage_a.pdf}
        \\[1em]
        \includegraphics[width=\linewidth]{figures/nullspace_preimage_b.pdf}
        \\[1em]
        \includegraphics[width=\linewidth]{figures/nullspace_preimage_c.pdf}
        \\[1em]
        \includegraphics[width=\linewidth]{figures/nullspace_preimage_d.pdf}
        \oscaption{nullspace_preimage}{%
            Same as Figures~\ref{fig:nullspace_preimage} and \ref{fig:nullspace_preimage_85}, \
            but assuming we can measure the light curves of these stars from many
            different inclinations. In this limit, the information content of our
            data approaches $50\%$ of the spatial information about the surface.
            \label{fig:nullspace_preimage_ensemble}
        }
    \end{centering}
\end{figure}

@@

\subsection{Ensemble analyses don't necessarily help}
\label{sec:basic-gp}
%
To explore this point in a bit more detail, let us consider the normalization
degeneracy in the context of ensemble analyses. Even though we can't
uniquely infer contrasts or numbers of spots from individual light curves,
perhaps we could harness the power of the ensemble.
%
Let us therefore go
back to our thought experiment in which we added spots to a stellar
surface. Assuming for simplicity that all spots have the same contrast $c$,
every time we add a spot the flux (in fractional units)
decreases by an amount proportional to $c$, so to first order we can approximate an
individual light curve as
%
\begin{linenomath}\begin{align}
        \label{eq:fapprox}
        \mathbf{f}(c, n) = \mathbf{1} - c \sum_{i=0}^{n-1} \mathbf{g}(\pmb{\theta}_i)
    \end{align}\end{linenomath}
%
where $\mathbf{g}(\pmb{\theta}_i)$ is some (complicated) function of the properties
of the $i^\mathrm{th}$ spot
as well as the stellar inclination, rotation period, etc., which we
denote by $\pmb{\theta}_i$.
Now, consider many stellar light curves drawn from some distribution
controlling the stellar and starspot properties with probability density
$p(\pmb{\theta})$.
%
The mean of the distribution of
light curves (still in fractional units) is then given by
%
\begin{linenomath}\begin{align}
        \mu(c, n) & = \mathrm{E} \Big[ \mathbf{f} (c, n) \Big] \nonumber                                      \\
                  & = 1 - c \, \mathrm{E} \bigg[ \sum_{i=0}^{n-1} \mathbf{g}(\pmb{\theta}_i) \bigg] \nonumber \\
                  & = 1 - c \, n \, \alpha
        \quad,
    \end{align}\end{linenomath}
%
where $\mathrm{E}\big[\cdots\big]$ denotes the expected value and
%
\begin{linenomath}\begin{align}
        \alpha \equiv \int \mathbf{g} (\pmb{\theta}) p(\pmb{\theta}) \mathrm{d} \pmb{\theta}
    \end{align}\end{linenomath}
%
is the expected value of $\mathbf{g}$.
%
Similarly, the variance of the distribution may be computed as
%
\begin{linenomath}\begin{align}
        \sigma^2(c, n) & = \mathrm{Var} \Big[ \mathbf{f}(c, n) \Big] \nonumber                         \nonumber \\
                       & = \mathrm{Var} \Big[ c \, \sum_{i=0}^{n-1} \mathbf{g}(\pmb{\theta}_i) \Big]   \nonumber \\
                       & = c^2 \, \mathrm{Var} \Big[ \sum_{i=0}^{n-1} \mathbf{g}(\pmb{\theta}_i) \Big] \nonumber \\
                       & = c^2 \sum_{i=0}^{n-1} \mathrm{Var} \Big[  \mathbf{g}(\pmb{\theta}_i) \Big]   \nonumber \\
                       & = c^2 \, n \, (\beta^2 - \alpha^2)
        \quad,
    \end{align}\end{linenomath}
%
where $\mathrm{Var}\big[\cdots\big]$ denotes the variance,
%
\begin{linenomath}\begin{align}
        \beta^2 \equiv \int \mathbf{g}^2 (\pmb{\theta}) p(\pmb{\theta}) \mathrm{d} \pmb{\theta}
    \end{align}\end{linenomath}
%
is the expected value of $\mathbf{g}^2$, and we used the fact that the variance
of the sum of independent random variables is equal to the sum of their variances.

@@

To summarize, the mean and variance of the ensemble of stellar light curves
in fractional units is
%
\begin{linenomath}\begin{align}
        \mu      & = 1 - c \, n \, \alpha
        \nonumber                                     \\
        \sigma^2 & = c^2 \, n \, (\beta^2 - \alpha^2)
        \quad,
    \end{align}\end{linenomath}
%
for some complicated functions $\alpha$ and $\beta$ of the distribution of
stellar inclinations,
rotation periods, and starspot properties.
%
If our observations were collected in these fractional units,
we could uniquely infer the spot contrast $c$ and the number of spots $n$,
since these enter as $c \, n$ and $c^2 n$ in the expressions for the mean
and variance, respectively, and these are straightforward statistics to
compute from the ensemble.
%
However, because of observations are made in \emph{relative} units, in which
we typically normalize to the mean, the amplitudes of features in light curves
can only tell us about the \emph{ratio}
%
\begin{linenomath}\begin{align}
        \label{eq:ratio}
        \frac{\sigma}{\mu}
         & \propto \frac{c \sqrt{n}}{1 - c n \alpha}
    \end{align}\end{linenomath}
%
for some value of $\alpha$.
In other words, even photometric ensemble analyses may not be able to tell
us about the values of the contrast and the number of spots independently.

A direct consequence of this normalization degeneracy is that it may not be
possible to uniquely constrain the total spot coverage of a star from
single-band photometry without strong prior assumptions.
The total spot coverage $f_S$
is simply the (average) area of a spot
times the total number of spots divided by the total area of the sphere, which
may be expressed as
%
\begin{linenomath}\begin{proof}{test_fS}
        \label{eq:fS}
        f_S = \frac{1}{2}\left(1 - \left<\cos r\right>\right)n
        \quad,
    \end{proof}\end{linenomath}
%
given an angular spot radius $r$.
While $r$ may be uniquely constrained from the covariance structure of the data
\citepalias{PaperII}, $n$ cannot.

@@

\begin{linenomath}\begin{proof}{test_decomposition}

@@

\begin{linenomath}\begin{proof}{test_PN}

@@

\begin{linenomath}\begin{proof}{test_orthogonality}

@@

Two things should be noted concerning this derivation. First,
because of the orthogonality of $\mathbf{V}$,
%
\begin{linenomath}\begin{proof}{test_PN}
        \mathbf{V} \mathbf{V}^\top &= \mathbf{I}
        \nonumber                                                 \\
        \mathbf{V}_\bullet \mathbf{V}_\bullet^\top  + \mathbf{V}_\circ \mathbf{V}_\circ^\top &= \mathbf{I}
        \nonumber \\
        \mathbf{P} + \mathbf{N} &= \mathbf{I}
        \quad,
    \end{proof}\end{linenomath}
%
so it follows from Equations~(\ref{eq:yrow}) and (\ref{eq:ynull})
that $\mathbf{y} = \mathbf{y}_\bullet + \mathbf{y}_\circ$.
